---
title: "Current AFL Men’s Predictions"
author: "James Day"
slug: current-afl-men-s-predictions
tags: []
categories: AFLM
---
Welcome to our all new AFL Men's ratings and tips page! Here you will the latest ratings, tips for the upcoming rounds and any simulations that have been done. I'll usually aim to update these early in the week to give you time to digest Any questions, let me know on Twitter! 


```{r setup, echo = FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)

round <-  2

HGA <- 36
B <- 0.025
k_val <- 18
carryOver <- 0.07
M <- 400
```

```{r get data, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
# Load libraries
library(fitzRoy)
library(tidyverse)
library(formattable)
library(ggthemes)
library(elo)

# Get fixture data using FitzRoy
fixture <- get_fixture()

# Get results
results <- fitzRoy::get_match_results()
results <- results %>%
  mutate(
    seas_rnd = paste0(Season, ".", Round.Number),
    First.Game = ifelse(Round.Number == 1, TRUE, FALSE)
  )

results_long <- convert_results(results)
```

## AFL Men's Ratings
Below are the updated AFL Men's ratings. The figure shows the current rating in green, with the tail indicating the previous rating for that team. 

```{r calculate ELO, echo=FALSE, message=FALSE, warning=FALSE}
# mapping
map_margin_to_outcome <- function(margin, B = 0.025) {
  outcome <- 1 / (1 + (exp(-B * margin)))
  return(outcome)
}

# mapping
map_outcome_to_margin <- function(outcome, B = 0.025) {
  margin <- log((1 / outcome) - 1) / -B
  return(margin)
}

  # calculate ELO
  elo.data <- elo.run(
    map_margin_to_outcome(Home.Points - Away.Points, B = B) ~
    adjust(Home.Team, HGA) +
      Away.Team +
      group(seas_rnd) +
      regress(First.Game, 1500, carryOver) +
      k(k_val * log(abs(Home.Points - Away.Points) + 1)),
    data = results
  )
```


```{r ratings, echo=FALSE, message=FALSE, warning=FALSE}

elo <- results %>%
  bind_cols(as.data.frame(elo.data)) %>%
  rename(Home.ELO = elo.A, 
         Away.ELO = elo.B) %>%
  mutate(Home.ELO_pre = Home.ELO - update,
         Away.ELO_pre = Away.ELO + update) %>%
  select(Date, Game, Home.Team, Away.Team, Home.ELO:Away.ELO_pre) 

elo_long <- elo %>%
  gather(variable, value, Home.Team:Away.ELO_pre) %>%
  separate(variable, into = c("Status", "variable"), sep = "\\.") %>%
  spread(variable, value) %>%
  mutate(ELO = as.numeric(ELO), ELO_pre = as.numeric(ELO_pre)) %>%
  group_by(Team) %>%
  arrange(Game) %>%
  filter(Game == last(Game) & !Team %in% c("Fitzroy", "University")) 
  
```

```{r ratings-plot, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(data = elo_long, aes(x = reorder(Team, ELO))) +
  geom_hline(yintercept = 1500, alpha = 0.7, linetype = 3) +
  geom_linerange(aes(ymin = ELO_pre, ymax = ELO), alpha = 0.3) +
  geom_point(aes(y = ELO_pre), alpha = 0.1) +
  geom_point(aes(y = ELO), colour = "#669999", size = 2) + 
  coord_flip() + 
  theme_minimal() +
  labs(x = "Team",
       y = "ELO Rating",
    title = "AFL Men's ELO Ratings",
       subtitle = "Showing the ELO rating of each AFL Men's team at the start of the 2018 
season in green, with the tail showing their end of 2017 rating",
       caption = "(data sourced from afltables.com)")

```


## Predictions

Below are the predictions for the current round. The margin and probability are for the home team winning, so a margin `< 0` or a probability `<0.5` indicates the away team is favourite. 

```{r Predictions, echo=FALSE, message=FALSE, warning=FALSE}
# Filter round and do predictions
predictions <- fixture %>%
  mutate(
    Day = format(Date, "%a, %d"),
    Time = format(Date, "%H:%M"),
    Probability = round(predict(elo.data, newdata = fixture), 3),
    Prediction = ceiling(map_outcome_to_margin(Probability)),
    Result = case_when(
           Probability > 0.5 ~ paste(Home.Team, "by", Prediction),
           Probability < 0.5 ~ paste(Away.Team, "by", -Prediction),
           TRUE ~ "Draw"
  )) %>%
  select(Day, Time, Round, Venue, Home.Team, Away.Team, Prediction, Probability, Result)


# Pass those onto Formattable
predictions %>%
  filter(Round == round) %>%
formattable(list(
  Home.Team = formatter("span", style = ~ style("font-weight" = ifelse(Probability >= 0.501, "bold", NA))),
  Away.Team = formatter("span", style = ~ style("font-weight" = ifelse(Probability <= 0.499, "bold", NA)))
  )
  ) 

```