---
title: 'Project: World Cup Datathon '
author: James Day
date: '2018-06-14'
slug: project-world-cup-datathon
categories:
  - World Cup Datathon
tags: []
subtitle: ''
---

In celebration of the World Cup in Russia, I've decided to take on Betfair's World Cup Datathon. See below my series of posts outlining how I built my models.


This is part of a series of posts on the World Cup Betfair datathon. See the links to others below.   
   
[Part 1 - Intro](https://plussixoneblog.com/post/football-world-cup-datathon-part-1)    
[Part 2 - Data Acquisition](https://plussixoneblog.com/post/football-world-cup-datathon-part-2)   
[Part 3 - Data Exploration and Feature Engineering](https://plussixoneblog.com/post/football-world-cup-datathon-part-3)    
Part 4 - Models (coming soon)   
Part 5 - Review (coming soon)   


# Predictions
The following are my actual predictions, based on each model. I'll play around with how to present them but you can see them below. 

```{r predict, echo=FALSE, message=FALSE, warning=FALSE}
library(pacman)
p_load(tidyverse, formattable, widgetframe)

# Load files
path <- here::here("projects", "world-cup-2018", "submissions")
files <- list.files(path)

dat <- paste0(path, "/", files) %>%
  map(read_csv) %>%
  map2_dfr(.y = files, ~ .x %>% mutate(model = str_remove(.y, "james_day_"),
                                       model = str_remove(model, ".csv"),
                                       group = str_extract(match_id, "[[a-z]]"),
                                       group_match = as.integer(str_extract(match_id, "\\d+"))))

unique(dat$model)
```

### Neural Net
```{r echo=FALSE, message=FALSE, warning=FALSE}
nnet <- dat %>% 
  filter(model == "NeuralNetwork") %>%
  formattable() %>%
  as.datatable(options = list(dom = 't', pageLength = 18))

frameWidget(nnet, height = 700, width = '95%')
```


### Multinomal
```{r echo=FALSE, message=FALSE, warning=FALSE}
multi <- dat %>% 
  filter(model == "PenalizedMultinomialRegressio") %>%
  formattable() %>%
  as.datatable(options = list(dom = 't', pageLength = 18))

frameWidget(multi, height = 700, width = '95%')
```

### GBM
```{r echo=FALSE, message=FALSE, warning=FALSE}
gbm <- dat %>% 
  filter(model == "StochasticGradientBoosting") %>%
  formattable() %>%
  as.datatable(options = list(dom = 't', pageLength = 18))

frameWidget(gbm, height = 700, width = '95%')
```

### Random Forest
```{r echo=FALSE, message=FALSE, warning=FALSE}
rf <- dat %>% 
  filter(model == "RandomForest") %>%
  formattable() %>%
  as.datatable(options = list(dom = 't', pageLength = 18))

frameWidget(rf, height = 700, width = '95%')
```

### CART
```{r echo=FALSE, message=FALSE, warning=FALSE}
cart <- dat %>% 
  filter(model == "CART") %>%
  formattable() %>%
  as.datatable(options = list(dom = 't', pageLength = 18))

frameWidget(cart, height = 700, width = '95%')
```

# Results
I'll post these after the fact.