---
title: Round 7 - AFL Mens Preview
author: James Day
date: '2018-05-04'
slug: round-7-afl-mens-preview
categories:
  - AFLM
tags: []
---


On the road again this weekend but I'll do my best to give a quick update! You can see my tips, as always, at [this link](http://plussixoneblog.com/page/aflm-current-tips/) or on [Squiggle](squiggle.com.au). 

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)

# Load libraries
library(pacman)
pacman::p_load(fitzRoy, tidyverse, formattable, ggthemes, elo, here, lubridate, widgetframe)

# Load data that has been run using 'weekly_data_process.R'
dat_list <- read_rds(here::here("data", "raw-data", "AFLM.rds"))
sim_dat_list <- read_rds(here::here("data", "raw-data", "AFLM_sims.rds"))
```

## Ratings
It's been a couple weeks since we've looked in depth at the ratings. In that time, as has widely been reported, Richmond have jumped out into clear air in the lead. They are clearly the strongest team in the competition at this point of the year. 

A pack of 4 is now chasing them in Sydney, Adelaide, GWS and Geelong. We've been lucky enough to see those teams fighting it out over the last few weeks with some suprising results and have another chance tonight in Geelong, with the Giants visiting. 

The next pack of 4 of West Coast, Hawthorn, Port Adelaide and, surprisingly, Collingwood. The Pies actually jumped up to 6th on our ratings last week but their big loss to Richmond (despite many reports of it being a positive loss) sees them lose a fair bit of ground. 

Melbourne and North Melbourne both hover around the average team mark, with then some clear steps back in rating points to our cellar dwellers. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Get the round data
elo_round <- dat_list$elo %>%
  filter(Game == last(Game) & !Team %in% c("Fitzroy", "University"))

sim_round <- sim_dat_list$sim_data_summary %>%
  filter(Round == last(dat_list$elo$Round.Number))

# Create a subtitle for plot
subt <- paste(
  "Showing the ELO rating of each AFL Men's team at the end of Round",
  dat_list$elo$Round.Number %>% last(), "",
  dat_list$elo$Date %>% last() %>% year(),
  "in green,\nwith the tail showing their previous rating"
)

# Do ggplot
elo_round %>%
  ggplot(aes(x = reorder(Team, ELO))) +
  geom_hline(yintercept = 1500, alpha = 0.7, linetype = 3) +
  geom_linerange(aes(ymin = ELO_pre, ymax = ELO), alpha = 0.3) +
  geom_point(aes(y = ELO_pre), alpha = 0.1) +
  geom_point(aes(y = ELO), colour = "#669999", size = 2) +
  coord_flip() +
  theme_minimal() +
  labs(
    x = "Team",
    y = "ELO Rating",
    title = "AFL Men's ELO Ratings",
    subtitle = subt,
    caption = "(data sourced from afltables.com)"
  )
```

## Simulations
Since we last talked, Richmond has firmed as minor premiership chances. They finish in top spot in 57% of all simulations - well above the next best in Sydney at 11%. The Tigers are also 2.5 games clear on our 'average wins' metric and appear to basically have finals and top 4 in the bag. 

Outside of those two teams, it is crazy tight at the top though. Seven teams (Adelaide, GWS, Hawks, Port, WCE, Geelong, Pies) have an average wins between 13.9 and 14.8! Each of those teams finish in the top 4 at least 30% of the time and have a greater than 70% chance of making finals. It's going to be facscinating to see that battle play out for finals spots. 

While the remaining teams are very long shots of getting a top 4 spot, we do have some teams with non-trivial chances of making finals. Melbourne, North and Freo all have >10% chance of reaching September. Outside of these teams, finals action looks unlikely. 


```{r echo=FALSE, message=FALSE, warning=FALSE}
# Combine Simulation and ELO data
elo_table <- elo_round %>%
  left_join(sim_round, by = "Team") %>%
  ungroup() %>%
  mutate(ELO.change = ELO - ELO_pre,
         change = min_rank(desc(ELO_pre)) - min_rank(desc(ELO))) %>%
  mutate_at(c("Top.8", "Top.4", "Top.1"), formattable::percent, digits = 1) %>%
  mutate_at(c("ELO", "ELO.change"), as.integer) %>%
  mutate_at("Wins", round, 1) %>%
  select(Team, change, ELO, ELO.change, Wins, Top.8, Top.4, Top.1) %>%
  arrange(desc(Wins))

# Write to formattable with some formatting
# Create a sign formatrer
sign_formatter <- formatter("span", 
                            style = x ~ style(color = case_when(
                              x > 0 ~ "green", 
                              x < 0 ~ "red",
                              TRUE ~ "transparent")))
# Create formattable/DT object
dt_elo <- elo_table %>%
  formattable(list(
    change = sign_formatter,
    ELO = normalize_bar("#20B2AA80"),
    ELO.change = sign_formatter,
    Top.8 = color_tile("transparent", "lightblue"),
    Top.4 = color_tile("transparent", "lightblue"),
    Top.1 = color_tile("transparent", "lightblue")
  )) %>% 
  as.datatable(options = list(dom = 't', pageLength = 18))

frameWidget(dt_elo, height = 700, width = '95%')
```

## Predictions
Onto this weeks games and it's looking a little lopsided at this stage. Tonight's game between Geelong and GWS sees two of our 'chasing 4' top teams face off. Geelong's strong home ground advantage looks to give them an edge. West Coast and Port is probably the only other game between teams in similar positions on our ratings - again, the home ground advantage looks to give the Eagles the edge here. 

Happy tipping and you can follow along with how these tips (and others) go at [The Squiggle](squiggle.com.au)

```{r predict, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
# Load libraries
library(pacman)
pacman::p_load(fitzRoy, tidyverse, formattable, 
               ggthemes, elo, here, lubridate, 
               widgetframe, ggridges, DT)

# Load data that has been run using 'weekly_data_process.R'
dat_list <- read_rds(here::here("data", "raw-data", "AFLM.rds"))

# Get predictions for current round and present as table
dat_list$predictions %>%
  filter(Round.Number == last(dat_list$elo$Round.Number + 1)) %>%
  select(-Time, -Round.Number) %>%
  mutate_at("Probability", formattable::percent, digits = 1) %>%
  formattable(list(
    Home.Team = formatter(
      "span", style =
        ~ style("font-weight" = ifelse(Probability >= 0.501, "bold", NA))
    ),
    Away.Team = formatter(
      "span", style =
        ~ style("font-weight" = ifelse(Probability <= 0.499, "bold", NA))
    )
  ))
```

