---
title: 'Round 4 AFL Men''s ratings '
author: James Day
date: '2018-04-09'
slug: round-4-afl-men-s-ratings
categories:
  - AFLM
tags: []
---
```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)

# Load libraries
library(pacman)
pacman::p_load(fitzRoy, tidyverse, formattable, ggthemes, elo, here, lubridate, widgetframe)

# Load data that has been run using 'weekly_data_process.R'
dat_list <- read_rds(here::here("data", "raw-data", "AFLM.rds"))
sim_dat_list <- read_rds(here::here("data", "raw-data", "AFLM_sims.rds"))
```

With round 3 done and dusted, are the top teams beginning to emerge? <!--more--> You can see the full ratings, results and tips [here](http://plussixoneblog.com/page/aflm-ratings-and-simulations/) but see our current ratings and simulations below. 


```{r echo=FALSE, message=FALSE, warning=FALSE}
# Get the round data
elo_round <- dat_list$elo %>%
  filter(Game == last(Game) & !Team %in% c("Fitzroy", "University"))

sim_round <- sim_dat_list$sim_data_summary %>%
  filter(Round == last(dat_list$elo$Round.Number))

# Create a subtitle for plot
subt <- paste(
  "Showing the ELO rating of each AFL Men's team at the end of Round",
  dat_list$elo$Round.Number %>% last(), "",
  dat_list$elo$Date %>% last() %>% year(),
  "in green,\nwith the tail showing their previous rating"
)

# Do ggplot
elo_round %>%
  ggplot(aes(x = reorder(Team, ELO))) +
  geom_hline(yintercept = 1500, alpha = 0.7, linetype = 3) +
  geom_linerange(aes(ymin = ELO_pre, ymax = ELO), alpha = 0.3) +
  geom_point(aes(y = ELO_pre), alpha = 0.1) +
  geom_point(aes(y = ELO), colour = "#669999", size = 2) +
  coord_flip() +
  theme_minimal() +
  labs(
    x = "Team",
    y = "ELO Rating",
    title = "AFL Men's ELO Ratings",
    subtitle = subt,
    caption = "(data sourced from afltables.com)"
  )
```



Not a whole lot of movement at the top, with Adelaide, Sydney, GWS, Richmond and Port remaining in the top tier of teams. These teams are all projected around 16 wins and ~80% chance of finals at this early stage, with Adelaide poking it's nose out in front. 

The pack following is lead by Geelong and rounded out by West Coast, Melbourne and Hawthorn. This cluster all sits at around ~13 projected wins and ~55% chance of finals. 

Of the rest, Collingwood looks the most likely to jump into that next tier, sitting just below the average rating of 1500 and a finals change of ~40%. Essendon, North and St Kilda fell away this round and now comfortably 'below average' teams, along with winners Freo and Bulldogs. 

It also looks like we can rule a line through Gold Coast, Brisbane and Carlton who are all projected below 7 wins for the year.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Combine Simulation and ELO data
elo_table <- elo_round %>%
  left_join(sim_round, by = "Team") %>%
  ungroup() %>%
  mutate(ELO.change = ELO - ELO_pre,
         change = min_rank(desc(ELO_pre)) - min_rank(desc(ELO))) %>%
  mutate_at(c("Top.8", "Top.4", "Top.1"), formattable::percent, digits = 1) %>%
  mutate_at(c("ELO", "ELO.change"), as.integer) %>%
  mutate_at("Wins", round, 1) %>%
  select(Team, change, ELO, ELO.change, Wins, Top.8, Top.4, Top.1) %>%
  arrange(desc(Wins))

# Write to formattable with some formatting
# Create a sign formatrer
sign_formatter <- formatter("span", 
                            style = x ~ style(color = case_when(
                              x > 0 ~ "green", 
                              x < 0 ~ "red",
                              TRUE ~ "transparent")))
# Create formattable/DT object
dt_elo <- elo_table %>%
  formattable(list(
    change = sign_formatter,
    ELO = normalize_bar("#20B2AA80"),
    ELO.change = sign_formatter,
    Top.8 = color_tile("transparent", "lightblue"),
    Top.4 = color_tile("transparent", "lightblue"),
    Top.1 = color_tile("transparent", "lightblue")
  )) %>% 
  as.datatable(options = list(dom = 't', pageLength = 18))

frameWidget(dt_elo, height = 700, width = '95%')

#change = formatter("span",
    #                   style = 
    #                     x ~ style(color = case_when(
    #                       x > 0 ~ "green",
    #                       x < 0 ~ "red",
    #                       TRUE ~ "grey")),
    #                   x ~ icontext(case_when(
    #                     x > 0 ~ "chevron-up",
    #                     x < 0 ~ "chevron-down",
    #                     TRUE ~ "minus")
    #                     )),
```


