---
title: Round 11 AFL Mens Preview
author: james day
date: '2018-06-01'
slug: round-11-afl-mens-preview
categories:
  - AFLM
tags: []
---

Round 11 is on us - read on for an overview of the weekend ahead. You can see my tips, as always, at [this link](http://plussixoneblog.com/page/aflm-current-tips/) or on [Squiggle](squiggle.com.au). 

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)

# Load libraries
library(pacman)
pacman::p_load(fitzRoy, tidyverse, formattable, ggthemes, elo, here, lubridate, widgetframe)

# Load data that has been run using 'weekly_data_process.R'
dat_list <- read_rds(here::here("data", "raw-data", "AFLM.rds"))
sim_dat_list <- read_rds(here::here("data", "raw-data", "AFLM_sims.rds"))
elo <- dat_list$elo
```

## Ratings
Don't look now but we have a new ELO leader! The Demons, on the back of three huge weeks have jumped the competition to be in top spot. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Get the round data
elo_round <- dat_list$elo %>%
  filter(Game == last(Game) & !Team %in% c("Fitzroy", "University"))

sim_round <- sim_dat_list$sim_data_summary %>%
  filter(Round == last(dat_list$elo$Round.Number))

# Create a subtitle for plot
subt <- paste(
  "Showing the ELO rating of each AFL Men's team at the end of Round",
  dat_list$elo$Round.Number %>% last(), "",
  dat_list$elo$Date %>% last() %>% year(),
  "in green,\nwith the tail showing their previous rating"
)

# Do ggplot
elo_round %>%
  ggplot(aes(x = reorder(Team, ELO))) +
  geom_hline(yintercept = 1500, alpha = 0.7, linetype = 3) +
  geom_linerange(aes(ymin = ELO_pre, ymax = ELO), alpha = 0.3) +
  geom_point(aes(y = ELO_pre), alpha = 0.1) +
  geom_point(aes(y = ELO), colour = "#669999", size = 2) +
  coord_flip() +
  theme_minimal() +
  labs(
    x = "Team",
    y = "ELO Rating",
    title = "AFL Men's ELO Ratings",
    subtitle = subt,
    caption = "(data sourced from afltables.com)"
  )
```

Given this, I wondered when was the last time they reached these heights? Firstly, we can look at the last time they had an ELO this high.

```{r}
elo %>%
  filter(Team == "Melbourne" & ELO >= last(ELO)) %>%
  arrange(desc(Date)) 
```

So they briefly were above their current level in 1994 and then not since wayyyy back in 1964, the year of their last Premiership. 

What about being on top of the ratings? 

```{r paged.print=TRUE}
elo %>%
  mutate(Season = year(Date)) %>%
  group_by(Season, Round) %>%
  mutate(rank = min_rank(desc(ELO))) %>%
  filter(Team == "Melbourne" & rank == 1) %>%
  arrange(desc(Date))

```

Again, we have to look to the 1994 season and then back to their premiership era. I wonder how long the lid will stay on? 

The remainder of our ratings reveals some distinct groups at the top - the top 4 of Melbourne, West Coast, Richmond and Sydney are all neck and neck. Geelong sits a bit in no-mans land between that top group and our chasing pack of finals aspirants in Adelaide, Port, Collingwood and North. 


## Simulations
Those top 9 teams continue to battle it out for our finals spots in our simulations - the leading pack all now but assured of making it with probabilities of >90%. Hawthorn, on the back of some banked wins and a presumably soft draw, still remains close to a 50% chance of making it, despite sitting a lowly 11th on our ratings. It is starting to look like a battle between them, North and Adelaide for that final 8th spot. 

Of course, as we've seen with GWS and Melbourne, those chances can change dramatically on the back of some big wins or losses! 

The race for the minor premiership has also been blown open - what was seemingly the Tigers to lose is now a pretty even race between West Coast, Richmond and Melbourne. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Combine Simulation and ELO data
elo_table <- elo_round %>%
  left_join(sim_round, by = "Team") %>%
  ungroup() %>%
  mutate(ELO.change = ELO - ELO_pre,
         change = min_rank(desc(ELO_pre)) - min_rank(desc(ELO))) %>%
  mutate_at(c("Top.8", "Top.4", "Top.1"), formattable::percent, digits = 1) %>%
  mutate_at(c("ELO", "ELO.change"), as.integer) %>%
  mutate_at("Wins", round, 1) %>%
  select(Team, change, ELO, ELO.change, Wins, Top.8, Top.4, Top.1) %>%
  arrange(desc(Wins))

# Write to formattable with some formatting
# Create a sign formatrer
sign_formatter <- formatter("span", 
                            style = x ~ style(color = case_when(
                              x > 0 ~ "green", 
                              x < 0 ~ "red",
                              TRUE ~ "transparent")))
# Create formattable/DT object
dt_elo <- elo_table %>%
  formattable(list(
    change = sign_formatter,
    ELO = normalize_bar("#20B2AA80"),
    ELO.change = sign_formatter,
    Top.8 = color_tile("transparent", "lightblue"),
    Top.4 = color_tile("transparent", "lightblue"),
    Top.1 = color_tile("transparent", "lightblue")
  )) %>% 
  as.datatable(options = list(dom = 't', pageLength = 18))

frameWidget(dt_elo, height = 700, width = '95%')
```

## Predictions
It looks like a pretty lopsided weekend on paper. The only game with a margin of less than 2 goals being the Port v Hawks game. Both teams are pretty precariously placed in our ratings and simulations so we may learn a bit about their finals chances! 

Happy tipping and you can follow along with how these tips (and others) go at [The Squiggle](squiggle.com.au)

```{r predict, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
# Load libraries
library(pacman)
pacman::p_load(fitzRoy, tidyverse, formattable, 
               ggthemes, elo, here, lubridate, 
               widgetframe, ggridges, DT)

# Load data that has been run using 'weekly_data_process.R'
dat_list <- read_rds(here::here("data", "raw-data", "AFLM.rds"))

# Get predictions for current round and present as table
dat_list$predictions %>%
  filter(Round.Number == last(dat_list$elo$Round.Number + 1)) %>%
  select(-Time, -Round.Number) %>%
  mutate_at("Probability", formattable::percent, digits = 1) %>%
  formattable(list(
    Home.Team = formatter(
      "span", style =
        ~ style("font-weight" = ifelse(Probability >= 0.501, "bold", NA))
    ),
    Away.Team = formatter(
      "span", style =
        ~ style("font-weight" = ifelse(Probability <= 0.499, "bold", NA))
    )
  ))
```

