---
title: Round 12 - AFL Mens Preview
author: James Day
date: '2018-06-08'
slug: round-12-afl-mens-preview
categories:
  - AFLM
tags: []
subtitle: ''
---

Super quick update this week sorry guys. You can see my tips, as always, at [this link](http://plussixoneblog.com/page/aflm-current-tips/) or on [Squiggle](squiggle.com.au). 

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)

# Load libraries
library(pacman)
pacman::p_load(fitzRoy, tidyverse, formattable, ggthemes, elo, here, lubridate, widgetframe)

# Load data that has been run using 'weekly_data_process.R'
dat_list <- read_rds(here::here("data", "raw-data", "AFLM.rds"))
sim_dat_list <- read_rds(here::here("data", "raw-data", "AFLM_sims.rds"))
elo <- dat_list$elo
```

## Ratings
Richmond take top spot back off Melbourne this week. The Dees do keep a handy gap to 3rd spot though, where we find it tough to split West Coast, Sydney and Geelong. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Get the round data
elo_round <- dat_list$elo %>%
  filter(Game == last(Game) & !Team %in% c("Fitzroy", "University"))

sim_round <- sim_dat_list$sim_data_summary %>%
  filter(Round == last(dat_list$elo$Round.Number))

# Create a subtitle for plot
subt <- paste(
  "Showing the ELO rating of each AFL Men's team at the end of Round",
  dat_list$elo$Round.Number %>% last(), "",
  dat_list$elo$Date %>% last() %>% year(),
  "in green,\nwith the tail showing their previous rating"
)

# Do ggplot
elo_round %>%
  ggplot(aes(x = reorder(Team, ELO))) +
  geom_hline(yintercept = 1500, alpha = 0.7, linetype = 3) +
  geom_linerange(aes(ymin = ELO_pre, ymax = ELO), alpha = 0.3) +
  geom_point(aes(y = ELO_pre), alpha = 0.1) +
  geom_point(aes(y = ELO), colour = "#669999", size = 2) +
  coord_flip() +
  theme_minimal() +
  labs(
    x = "Team",
    y = "ELO Rating",
    title = "AFL Men's ELO Ratings",
    subtitle = subt,
    caption = "(data sourced from afltables.com)"
  )
```

## Simulations
Richmond have firmed back ahead of West Coast as our favourite for the minor premiership. The only other realistic show here is Melbourne, who are actually more likely now than West Coast - they top the ladder in 23% of simulations! 

If we take 20% as our cutoff for realistic finals chances this far out - we still have 11 teams fighting it out, just! GWS is rapidly slipping although their win on the weekend no doubt helped. Adelaide is the surprise here. Last seasons grand finalists now sit at 10th on our simulated wins ladder and only make finals in 32% of sims. There is a bunch of teams all pretty even between 50% and 80%. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Combine Simulation and ELO data
elo_table <- elo_round %>%
  left_join(sim_round, by = "Team") %>%
  ungroup() %>%
  mutate(ELO.change = ELO - ELO_pre,
         change = min_rank(desc(ELO_pre)) - min_rank(desc(ELO))) %>%
  mutate_at(c("Top.8", "Top.4", "Top.1"), formattable::percent, digits = 1) %>%
  mutate_at(c("ELO", "ELO.change"), as.integer) %>%
  mutate_at("Wins", round, 1) %>%
  select(Team, change, ELO, ELO.change, Wins, Top.8, Top.4, Top.1) %>%
  arrange(desc(Wins))

# Write to formattable with some formatting
# Create a sign formatrer
sign_formatter <- formatter("span", 
                            style = x ~ style(color = case_when(
                              x > 0 ~ "green", 
                              x < 0 ~ "red",
                              TRUE ~ "transparent")))
# Create formattable/DT object
dt_elo <- elo_table %>%
  formattable(list(
    change = sign_formatter,
    ELO = normalize_bar("#20B2AA80"),
    ELO.change = sign_formatter,
    Top.8 = color_tile("transparent", "lightblue"),
    Top.4 = color_tile("transparent", "lightblue"),
    Top.1 = color_tile("transparent", "lightblue")
  )) %>% 
  as.datatable(options = list(dom = 't', pageLength = 18))

frameWidget(dt_elo, height = 700, width = '95%')
```

## Predictions
A few intersting matchups in the shortened bye round. Friday's game looks like a cracker - I noticed on [The Squiggle](squiggle.com.au) that Port were actually the bookie favourites despite all the models tipping Richmond. Port is one of those teams in the logjam of 5th to 11th so a good showing is important. 

Bookending the weekend is the Queens Birthday clash on Monday. This might be the first one in a long time that Melbourne go in as favourites? Or at least the first one where both teams are actually pretty highly rated. Again, Collingwood is in that logjam in the middle so a win will be pretty big for them. 

Happy tipping and you can follow along with how these tips (and others) go at [The Squiggle](squiggle.com.au)

```{r predict, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
# Load libraries
library(pacman)
pacman::p_load(fitzRoy, tidyverse, formattable, 
               ggthemes, elo, here, lubridate, 
               widgetframe, ggridges, DT)

# Load data that has been run using 'weekly_data_process.R'
dat_list <- read_rds(here::here("data", "raw-data", "AFLM.rds"))

# Get predictions for current round and present as table
dat_list$predictions %>%
  filter(Round.Number == last(dat_list$elo$Round.Number + 1)) %>%
  select(-Time, -Round.Number) %>%
  mutate_at("Probability", formattable::percent, digits = 1) %>%
  formattable(list(
    Home.Team = formatter(
      "span", style =
        ~ style("font-weight" = ifelse(Probability >= 0.501, "bold", NA))
    ),
    Away.Team = formatter(
      "span", style =
        ~ style("font-weight" = ifelse(Probability <= 0.499, "bold", NA))
    )
  ))
```


