---
title: Round 5 - AFL Men’s ratings updates and preview
author: James Day
date: '2018-04-20'
slug: round-5-afl-men-s-ratings-updates-and-preview
categories:
  - AFLM
  - Predictions
tags: []
---
```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)

# Load libraries
library(pacman)
pacman::p_load(fitzRoy, tidyverse, formattable, ggthemes, elo, here, lubridate, widgetframe)

# Load data that has been run using 'weekly_data_process.R'
dat_list <- read_rds(here::here("data", "raw-data", "AFLM.rds"))
sim_dat_list <- read_rds(here::here("data", "raw-data", "AFLM_sims.rds"))
```

After a pretty brutal Round 4, we saw some big movements in our ratings. That leads to some interesting matchups this week. Read on to see who our model is tipping! 

<!--more-->

Also a quick reminder that you always find [full ratings](http://plussixoneblog.com/page/aflm-ratings-and-simulations/) and [current tips](http://plussixoneblog.com/page/aflm-current-tips/) updated each Monday. 

## Ratings and Simulations

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Get the round data
elo_round <- dat_list$elo %>%
  filter(Game == last(Game) & !Team %in% c("Fitzroy", "University"))

sim_round <- sim_dat_list$sim_data_summary %>%
  filter(Round == last(dat_list$elo$Round.Number))

# Create a subtitle for plot
subt <- paste(
  "Showing the ELO rating of each AFL Men's team at the end of Round",
  dat_list$elo$Round.Number %>% last(), "",
  dat_list$elo$Date %>% last() %>% year(),
  "in green,\nwith the tail showing their previous rating"
)

# Do ggplot
elo_round %>%
  ggplot(aes(x = reorder(Team, ELO))) +
  geom_hline(yintercept = 1500, alpha = 0.7, linetype = 3) +
  geom_linerange(aes(ymin = ELO_pre, ymax = ELO), alpha = 0.3) +
  geom_point(aes(y = ELO_pre), alpha = 0.1) +
  geom_point(aes(y = ELO), colour = "#669999", size = 2) +
  coord_flip() +
  theme_minimal() +
  labs(
    x = "Team",
    y = "ELO Rating",
    title = "AFL Men's ELO Ratings",
    subtitle = subt,
    caption = "(data sourced from afltables.com)"
  )
```

The big loser here was clearly Adelaide, after a shock, heavy loss to Collingwood. They dropped from pretty clear top rated team down to 4th. Their Finals chances took a big hit, dropping down to 8th on our average simulated season ladder with a finals chance now below 70%. 

Richmond's big win vaulted them up to top spot on our ratings and also our simulated season! They finish top spot in nearly 30% of all (albeit very early) simulations. 

Ratings wise, there is now a pretty linear progression from 1st down to 9th, with maybe two groupings of top 4 (Richmond, Sydney, GWS, Adelaide) and 5th-9th (Geelong, West Coast, Port Adelaide, Hawthorn, Collingwood). These top 9 are, as expected, also finishing making finals in most simulations. All have >60% chance of reaching September.  

There is a pretty big drop off after that to three even teams in North Melbourne, Melbourne and Essendon, all sitting just below the average rated team. They aren't out of finals by a long shot but only make it that far in about 30% of simulations. 

There is another drop to the Dogs, Saints and Dockers group before we have a decent step back to our cellar dwellers in Brisbane, Gold Coast and Carlton. All of those teams would have to launch a pretty epic turnaround to make it even from this early stage you'd think.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Combine Simulation and ELO data
elo_table <- elo_round %>%
  left_join(sim_round, by = "Team") %>%
  ungroup() %>%
  mutate(ELO.change = ELO - ELO_pre,
         change = min_rank(desc(ELO_pre)) - min_rank(desc(ELO))) %>%
  mutate_at(c("Top.8", "Top.4", "Top.1"), formattable::percent, digits = 1) %>%
  mutate_at(c("ELO", "ELO.change"), as.integer) %>%
  mutate_at("Wins", round, 1) %>%
  select(Team, change, ELO, ELO.change, Wins, Top.8, Top.4, Top.1) %>%
  arrange(desc(Wins))

# Write to formattable with some formatting
# Create a sign formatrer
sign_formatter <- formatter("span", 
                            style = x ~ style(color = case_when(
                              x > 0 ~ "green", 
                              x < 0 ~ "red",
                              TRUE ~ "transparent")))
# Create formattable/DT object
dt_elo <- elo_table %>%
  formattable(list(
    change = sign_formatter,
    ELO = normalize_bar("#20B2AA80"),
    ELO.change = sign_formatter,
    Top.8 = color_tile("transparent", "lightblue"),
    Top.4 = color_tile("transparent", "lightblue"),
    Top.1 = color_tile("transparent", "lightblue")
  )) %>% 
  as.datatable(options = list(dom = 't', pageLength = 18))

frameWidget(dt_elo, height = 700, width = '95%')
```

## Tips
Onto the tips for this week. 

```{r}
# Get predictions for current round and present as table
dat_list$predictions %>%
  filter(Round.Number == last(dat_list$elo$Round.Number + 1)) %>%
  select(-Time, -Round.Number) %>%
  mutate_at("Probability", formattable::percent, digits = 1) %>%
  formattable(list(
    Home.Team = formatter(
      "span", style =
        ~ style("font-weight" = ifelse(Probability >= 0.501, "bold", NA))
    ),
    Away.Team = formatter(
      "span", style =
        ~ style("font-weight" = ifelse(Probability <= 0.499, "bold", NA))
    )
  ))

```

## Other business
### Squiggle
I encourage you all to follow the Squiggle on [Twitter](https://twitter.com/SquiggleAFL/) and at the [website](sqiggle.com.au). It collates tips from a range of different models. Max is putting a lot of great work in this year and the list of models is growing weekly! 

### fitzRoy
Over the summer, [Rob](https://twitter.com/roberto_boberto/) of [Analysis of AFL](https://anoafl.github.io/) fame and I started working on an R package for AFL. It's called fitzRoy and is in it's early stages. If you are into that kind of thing, you can see instructions for using it [here](https://jimmyday12.github.io/fitzRoy/). Any feedback, suggestions or contributions are more than welcome! 

### New Logo
You'll notice a new Koala logo. After 2 years of not having any kind of logo, I decided (thanks to Squiggle) that I needed something. My good friend and colleague, Tina Hsu (she doesn't have Twitter) made me one. Many thanks to her! 

### AFL Wonk
I've started a [Slack group](slack.com) for discussing AFL analytic's. There is a great, diverse group of people so far in there. We've got some plans to do regular questions, sharing of data, ideas and general chit chat. If you'd like to join, it's open to anyone. I do however need an email to sign you up - if you could fill that out [here](https://goo.gl/forms/4rNDx8bDyck4itLu1) I'll set you up within a day or so. I promise no spamming (I've no idea how to do that anyway).
