---
title: Squiggle Disagreement
author: James Day
date: '2018-05-01'
slug: squiggle-disagreement
categories:
  - Squiggle
tags: []
draft: true
---

For those who follow me, I'm sure you are well aware of [The Squiggle](squiggle.com.au) run by Max Barry. In recent years, Max has done some amazing work at combining a bunch of different Footy models in one place. He recently created an API and [Rob](LINK) and I have now integrated it into the latest version of [`fitzRoy`](LINK).  

There is a lot of value at having this information available. I'm hoping to delve into that throughout the year.

This post is my first dive into it and looks into how much agreement there is between our models. 

## Getting the data
First of all, we can get the data with `fitzRoy`. 

```{r}
library(fitzRoy)
library(tidyverse)
results <- get_squiggle_data("tips")
glimpse(results)
```
We need our data in wide format to use the `dist` function. 

```{r}
results_w <- results %>%
  select(source, gameid, margin) %>%
  spread(gameid, margin)

dist(results_w)

```
