---
title: 'World Cup Datathon - Part 3: Feature Engineering'
author: James Day
date: '2018-06-12'
slug: world-cup-datathon-part-3
categories:
  - World Cup Datathon
tags: []
subtitle: ''
draft: true
---

In [Part 1](https://plussixoneblog.com/post/football-world-cup-datathon-part-1) and  
[Part 2](https://plussixoneblog.com/post/football-world-cup-datathon-part-1) of this series, I introduced the Betfair World Cup datathon and aquired some data. Now, let's spend a bit of time exploring that data and creating some features! 

Feature Engineering is where the art comes into our modelling process. Given I don't have a lot of domain knowledge here, I've done a bit of quick reading. Again - I'm not going to go into some advanced Soccer analystics like [Expected Goal (XG)](https://understat.com/) or any kind of network analysis - so I'll simply use the match results, FIFA rankings and betfair data. 

First - let's load our combined data from [Part 2](https://plussixoneblog.com/post/football-world-cup-datathon-part-1).

```{r setup, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

```{r Load}
# Load libraries
library(pacman)
pacman::p_load(tidyverse, here, elo, ggthemes, zoo, nanier, visdat, lubridate)

dat <- read_csv(here::here("projects", "world-cup-2018", "combined-data.csv"))
world_cup <- read_csv(here::here("projects", "world-cup-2018", "world-cup.csv"))
```

# Data Exploration
Let's do a quick exploration of our `dat` dataframe. I did a little it of this in [Part 2](https://plussixoneblog.com/post/football-world-cup-datathon-part-1) but we'll try get a bit more in depth.

Before we start, let's create our output variable and then look at it's distribution. 

```{r base}
dat <- dat %>%
  mutate(
    result = case_when(
      team_1_goals > team_2_goals ~ "win",
      team_1_goals < team_2_goals ~ "lose",
      TRUE ~ "draw"),
    result = as.factor(result))

dat %>%
  ungroup() %>%
  count(result) %>%
  mutate(perc = n/sum(n))
```
So - that is essentially my baseline model. For an game I can naively put in those percentages. If any other model can't beat that, it's not adding anything! 

What about getting an understanding of the rest of the data. 

```{r data_exp1}
glimpse(dat)
```

The first thing that jumps out at me is the `tournament` column contains the tournament and the year in it. It would be nice to group say, all "Euro" tournements together. I also want to get a match_id that is just the game number in our datset. 

```{r data_exp2}
dat <- dat %>%
  arrange(date) %>%
  mutate(tournament_cat = as.factor(str_remove(tournament, "[[:digit:]]+") %>% str_squish()),
         match_id = row_number())

dat %>% distinct(tournament_cat)
```

That's better. The other thing is that the `odds` data is being read as characters rather than numneric. That's pretty easy to fix. 


```{r data_exp3}
dat <- dat %>%
  mutate_at(vars(contains("odds")), as.numeric)
```

Now we can use `skim` to look at how each column is distributed or structured. 

```{r data_exp4}
skimr::skim(dat)
```



First thing to note is that we have a large chunk of our odds data missing. There is a smaller group of fifa ratings missing and then after that, all our data is complete. We can also see from that plot that a lot of our numeric variables are skewed in their distribution. In particular, the betfair odds seem to have some really large outliers at the top end. 


```{r visdat}
visdat::vis_dat(dat)
```

This plot shows us where the missing data occurs. You can see that a lot of the missing betfair data is also missing from the other odds data. One suspects that these are probably either early games in our dataset or from pretty small competitions. The FIFA data doesn't seem to have that same relationship.

Firstly, let's just remove the `odds` columns - that's way too many missing variables unfortunately. 

```{r remove_missing}
dat <- dat %>%
  select(-avg_odds_draw, -avg_odds_away_win, -avg_odds_home_win)
```

Let's see if we can find if the betfair or Fifa data that is missing has something common. 

```{r miss_fct}
naniar::gg_miss_fct(dat, tournament_cat)
```
So our `fifa` data is mostly missing for World Cups. I suspect that this has something to do with when those ratings are done - typically out of season. Maybe the FIFA ratings aren't updated as often enough.

```{r fifa_fix1}
dat %>%
  filter(tournament_cat == "World Cup") %>%
  select(date, team_1, team_2, team_1_fifa, team_2_fifa)

```

Let's pick a date and take a look. I'm going to look through the 2010 World Cup for Australia. 

```{r fifa_fix2}
dat %>%
  filter((team_1 == "Australia" | team_2 == "Australia") & year == 2010) %>%
  select(date, team_1, team_2, team_1_fifa, team_2_fifa, tournament)

fifa_dat <- read_csv(here::here("projects", "world-cup-2018", "fifa_rank_history.csv"))
fifa_dat %>%
  filter(Team == "Australia" & year(Date)  == 2010)
```

So - there was no rating update during the month of the World Cup. I think the best bet here is to simply move the data forward (i.e. take the rating from the month beforehand). That seems like a safe thing to do in general here. 

```{r fifa_fix3}
# Make fifa data long again and then fill
fifa <- dat %>%
  gather(team, team_val, team_1:team_2) %>%
  select(date, match_id, team, team_val, team_1_fifa, team_2_fifa) %>%
  mutate(fifa_points = ifelse(team == "team_1", team_1_fifa, team_2_fifa)) %>%
  group_by(team_val) %>%
  arrange(team_val, date) %>%
  fill(fifa_points, .direction = "down") %>%
  fill(fifa_points, .direction = "up") %>%
  select(-team_1_fifa, -team_2_fifa, -team)

# Join back in with dat
dat <- dat %>%
  select(-team_1_fifa, -team_2_fifa) %>%
  left_join(fifa, by = c("date", "match_id", "team_1" = "team_val")) %>%
  rename(team_1_fifa = fifa_points) %>%
  left_join(fifa, by = c("date", "match_id","team_2" = "team_val")) %>%
  rename(team_2_fifa = fifa_points)
  
summary(dat)
```

So we've fixed up most of the fifa data.

Onto the betfair data. It looked like the most missing data was coming from Oceania Nations Cup, Friendlies and African Cup of Nations. 

What about over time? We'd suspect older games to be less likely to have odds data. 

```{r missing_1}
dat %>%
  ggplot(aes(x = date, fill = is.na(team_1_betfair_odds))) +
  geom_histogram(binwidth = 30) + 
  theme_minimal()
```

Yep! So - the Betfair data isn't just randomly missing. It is mostly likely to be missing for older data and most likely to be missing for certain less popular tournamnts and friendlies. I'll probably remove that data later but I need those games to do some feature engineering, like ELO, so I'll leave it for now.

# Feature Engineering
## ELO
For my AFL ELO ratings, I've recently started using a fantastic R package called `elo`. You can find details of that package [here](https://github.com/eheinzen/elo) and my weekly R code [here](https://github.com/jimmyday12/plussixoneblog/blob/master/scripts/weekly_data_process.R). Again - not knowing how to best implement an ELO model in Football specifically, I've left it pretty simple. I'm following loosely the method employed at [eloratings.net](http://www.eloratings.net/about).

Firstly, I need to clasify the tournaments into groupings so that I can assign different k values for them. Namely,

    60 for World Cup finals;
    50 for continental championship finals and major intercontinental tournaments;
    40 for World Cup and continental qualifiers and major tournaments;
    30 for all other tournaments;
    20 for friendly matches.
    
I also need to adjust the k value based on the goal difference and give a 100 point HGA to any team playing at home. 

```{r elo_calcs}
# Function to clasify tournament
find_k <- function(x){
  case_when(
    x == "World Cup" ~ 60,
    x %in% c("Euro", "Copa America", "African Cup of Nations", 
             "Asian Cup", "Gold Cup", "Confederations Cup") ~ 50,
    str_detect(x, "Qualifiers") ~ 40,
    x == "Friendly" ~ 20,
    TRUE ~ 30
  )
}

# Function to adjust based on goal difference
k_adjust <- function(x){
  case_when(
    x == 2 ~ 1.5,
    x == 3 ~ 1.75,
    x > 3 ~ 1.75 + ((x - 3)/8),
    TRUE ~ 1
  )
}
```

Now, we can go and apply those functions to our data and perform the `elo.run` function. 

```{r elo}
dat <- dat %>%
  ungroup() %>%
  mutate(game_id = row_number()) %>%
  group_by(team_1) %>%
  mutate(
    team_1_result = case_when(
      team_1_goals < team_2_goals ~ 0,
      team_1_goals > team_2_goals ~ 1,
      TRUE ~ 0.5),
    team_2_result = 1 - team_1_result,
    team_1_goals_against = team_2_goals,
    team_2_goals_against = team_1_goals)

# Clasify tournemt and calculate k value
dat <- dat %>%
  mutate(
         k_val = find_k(tournament_cat),
         hga = ifelse(is_team_1_home, 100, 0)) %>%
  arrange(date)

# Run ELO
elo.data <- elo.run(
  team_1_result ~
    adjust(team_1, hga) + 
    team_2  +
    k(k_adjust(abs(team_1_goals - team_2_goals)) * k_val),
  data = dat
)

dat <- dat %>%
  bind_cols(as.data.frame(elo.data)) %>%
  mutate(team_1_elo = elo.A - update,
         team_2_elo = elo.B + update, 
         team_1_elo_prob = p.A) %>%
  select(date:k_val, team_1_elo:team_1_elo_prob)

```

Let's do a quick exploration of those ELO values.

```{r elo_viz}
dat %>%
  ggplot(aes(team_1_elo)) +
  geom_histogram()

final.elos(elo.data) %>%
  data.frame(elo = .) %>%
  rownames_to_column("Team") %>%
  filter(Team %in% world_cup$team_1) %>%
  arrange(desc(elo))
```

## Team history
For these next few features, I actually need to make my dataset 'long'. This allows me to calculate some historical information such as the number of games or wins over a period of team, if they won a cup or other features that might be interesting.

```{r team_long, message=FALSE, warning=FALSE}
team_dat <- dat %>%
  ungroup() %>%
  select(date, game_id, tournament_cat, contains("team")) %>%
  gather(variable, value, -date, -game_id, -tournament_cat) %>%
  separate(variable, into = c("team", "number", "measure"), extra = "merge") %>%
  mutate(measure = ifelse(is.na(measure), "team", measure)) %>%
  spread(measure, value) %>%
  select(date, game_id, tournament_cat, team, result, goals, goals_against) %>%
  na.omit()

head(team_dat)
nrow(team_dat)
nrow(dat) * 2
```
Ok - that was tricky. Oh well - we got it! 

## Values for Last 10 games
One thing that I've seen in a few prediction models is looking at the results and scores for a team over their last n games. In theory, some of this should be reflected in our ELO rating. But nonetheless, our ELO rating isn't a very sophisticated one (it looks at just results rather than scores), so I can add this in. Since I'm going to put it into some machine learning models, we can trim these out if they aren't very useful.

I've picked the last 10 games based on a couple examples I found [here](https://github.com/mrthlinh/FIFA-World-Cup-Prediction) and 

First - it makes sense to see how many games a team has won in it's last 10 games. I'll do this as their average value of result where a win is assigned 1, a draw is 0.5 and a loss is 0. I'll also calculate the average goals for and against in that period. 

```{r rolling, message=FALSE, warning=FALSE}
# Specify variables
variables <- vars(goals, result, goals_against)
team_dat <- team_dat %>%
  group_by(team) %>%
  arrange(team, date) %>%
  mutate_at(variables, as.numeric) %>%
  mutate(result_game = result) %>%
  mutate_at(variables,
            .funs = rollmean, k = 10, na.rm = T, fill = NA, partial = T, align = "right") %>%
  rename_at(variables, funs(paste0("last_10_", .)))
```

## World Cup Appearances
Being at a World Cup multiple times generally, you'd think, would be a good predictor of performing well at world cups. In particular, playing more games would seem to indicate that you've done well at world cups. 

```{r world_cup}
team_dat <- team_dat %>%
  group_by(team) %>%
  mutate(world_cup_wins = cumsum(tournament_cat == "World Cup" & result_game == 1),
         world_cup_games = cumsum(tournament_cat == "World Cup")) %>%
  select(-tournament_cat, -result_game, -date)

dat <- dat %>%
  left_join(team_dat, by = c("game_id", "team_1" = "team")) %>%
  left_join(team_dat, by = c("game_id", "team_2" = "team"), suffix = c("_team_1", "_team_2"))
```


# Test Data Set
For our final data, we need all of these variables to exist. 

```{r world_cup_add}
# Add ELO
final_elo <- final.elos(elo.data) %>%
  data.frame(elo = .) %>%
  rownames_to_column("Team")

world_cup <- world_cup %>%
  left_join(final_elo, by = c("team_1" = "Team")) %>%
  rename(team_1_elo = elo) %>%
  left_join(final_elo, by = c("team_2" = "Team")) %>%
  rename(team_2_elo = elo)
  
# Add the team history data - take last value
team_dat_last <- team_dat %>%
  group_by(team) %>%
  filter(row_number() == n())

world_cup <- world_cup %>%
  left_join(team_dat_last, by = c("team_1" = "team")) %>%
  left_join(team_dat_last, by = c("team_2" = "team"), suffix = c("_team_1", "_team_2"))
```

# Save
Now we can save our data.

```{r save, message=FALSE, warning=FALSE}
write_csv(dat, here::here("projects", "world-cup-2018", "combined-data-cleaned.csv"))
write_csv(world_cup, here::here("projects", "world-cup-2018", "world-cup-cleaned.csv"))
```

---
This is part of a series of posts on the World Cup Betfair datathon. See the links to others below.   
   
[Part 1 - Intro](https://plussixoneblog.com/post/football-world-cup-datathon-part-1)    
[Part 2 - Data Acquisition](https://plussixoneblog.com/post/football-world-cup-datathon-part-1)   
Part 3 - Data Exploration and Feature Engineering (coming soon)   
Part 4 - Models (coming soon) 
Part 5 - Review (coming soon)
---
