---
title: AFLM - Round 13 preview
author: James Day
date: '2018-06-14'
slug: aflm-round-13-preview
categories:
  - AFLM
tags: []
subtitle: ''
---

Those who have been following the blog for a while know that Thursday games always catch me out and this week is no exception! No update for me but you can see my tips, as always, at [this link](http://plussixoneblog.com/page/aflm-current-tips/) or on [Squiggle](squiggle.com.au). 

I've also got my plots below, presented without comment. 

As a separate note - I've decided to enter the World Cup Datathon. You can see my posts below More updates coming. 


----
[Part 1 - Intro](https://plussixoneblog.com/post/football-world-cup-datathon-part-1)    
[Part 2 - Data Acquisition](https://plussixoneblog.com/post/football-world-cup-datathon-part-2)   
[Part 3 - Data Exploration and Feature Engineering]((https://plussixoneblog.com/post/football-world-cup-datathon-part-3)     
Part 4 - Models (coming soon)   
Part 5 - Review (coming soon)    

----


```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)

# Load libraries
library(pacman)
pacman::p_load(fitzRoy, tidyverse, formattable, ggthemes, elo, here, lubridate, widgetframe)

# Load data that has been run using 'weekly_data_process.R'
dat_list <- read_rds(here::here("data", "raw-data", "AFLM.rds"))
sim_dat_list <- read_rds(here::here("data", "raw-data", "AFLM_sims.rds"))
elo <- dat_list$elo
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
# Get the round data
elo_round <- dat_list$elo %>%
  filter(Game == last(Game) & !Team %in% c("Fitzroy", "University"))

sim_round <- sim_dat_list$sim_data_summary %>%
  filter(Round == last(dat_list$elo$Round.Number))

# Create a subtitle for plot
subt <- paste(
  "Showing the ELO rating of each AFL Men's team at the end of Round",
  dat_list$elo$Round.Number %>% last(), "",
  dat_list$elo$Date %>% last() %>% year(),
  "in green,\nwith the tail showing their previous rating"
)

# Do ggplot
elo_round %>%
  ggplot(aes(x = reorder(Team, ELO))) +
  geom_hline(yintercept = 1500, alpha = 0.7, linetype = 3) +
  geom_linerange(aes(ymin = ELO_pre, ymax = ELO), alpha = 0.3) +
  geom_point(aes(y = ELO_pre), alpha = 0.1) +
  geom_point(aes(y = ELO), colour = "#669999", size = 2) +
  coord_flip() +
  theme_minimal() +
  labs(
    x = "Team",
    y = "ELO Rating",
    title = "AFL Men's ELO Ratings",
    subtitle = subt,
    caption = "(data sourced from afltables.com)"
  )
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
# Combine Simulation and ELO data
elo_table <- elo_round %>%
  left_join(sim_round, by = "Team") %>%
  ungroup() %>%
  mutate(ELO.change = ELO - ELO_pre,
         change = min_rank(desc(ELO_pre)) - min_rank(desc(ELO))) %>%
  mutate_at(c("Top.8", "Top.4", "Top.1"), formattable::percent, digits = 1) %>%
  mutate_at(c("ELO", "ELO.change"), as.integer) %>%
  mutate_at("Wins", round, 1) %>%
  select(Team, change, ELO, ELO.change, Wins, Top.8, Top.4, Top.1) %>%
  arrange(desc(Wins))

# Write to formattable with some formatting
# Create a sign formatrer
sign_formatter <- formatter("span", 
                            style = x ~ style(color = case_when(
                              x > 0 ~ "green", 
                              x < 0 ~ "red",
                              TRUE ~ "transparent")))
# Create formattable/DT object
dt_elo <- elo_table %>%
  formattable(list(
    change = sign_formatter,
    ELO = normalize_bar("#20B2AA80"),
    ELO.change = sign_formatter,
    Top.8 = color_tile("transparent", "lightblue"),
    Top.4 = color_tile("transparent", "lightblue"),
    Top.1 = color_tile("transparent", "lightblue")
  )) %>% 
  as.datatable(options = list(dom = 't', pageLength = 18))

frameWidget(dt_elo, height = 700, width = '95%')
```

```{r predict, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
# Load libraries
library(pacman)
pacman::p_load(fitzRoy, tidyverse, formattable, 
               ggthemes, elo, here, lubridate, 
               widgetframe, ggridges, DT)

# Load data that has been run using 'weekly_data_process.R'
dat_list <- read_rds(here::here("data", "raw-data", "AFLM.rds"))

# Get predictions for current round and present as table
dat_list$predictions %>%
  filter(Round.Number == last(dat_list$elo$Round.Number + 1)) %>%
  select(-Time, -Round.Number) %>%
  mutate_at("Probability", formattable::percent, digits = 1) %>%
  formattable(list(
    Home.Team = formatter(
      "span", style =
        ~ style("font-weight" = ifelse(Probability >= 0.501, "bold", NA))
    ),
    Away.Team = formatter(
      "span", style =
        ~ style("font-weight" = ifelse(Probability <= 0.499, "bold", NA))
    )
  ))
```


