---
title: Round 11 AFLM Predictions
author: James Day
date: '2019-05-27'
slug: round-11-aflm-predictions
categories:
  - AFLM
tags: []
subtitle: ''
---

Hi team,

See the tips for *round 11* of the AFL Men's comp below, as well as some simulation results. 

You can always receive my live, up to date tips [here](https://plussixoneblog.com/page/aflm-current-tips/) and my simulations  [here](https://plussixoneblog.com/page/aflm-ratings-and-simulations/) regardless of the timing or existance of this weekly post! 

And reach out on [Twitter](https://twitter.com/plusSixOneblog) if you have any thoughts or questions.

Enjoy and happy tipping :) 

## Tips

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

```{r}
# Load libraries
library(pacman)
pacman::p_load(fitzRoy, tidyverse, formattable, 
               ggthemes, elo, here, lubridate, 
               widgetframe, ggridges, DT)

round_1 <- FALSE

# Load data that has been run using 'weekly_data_process.R'
dat_list <- read_rds(here::here("data", "raw-data", "AFLM.rds"))

round_num <- ifelse(round_1, 1, last(dat_list$elo$Round.Number + 1))

# Get predictions for current round and present as table
dat_list$predictions %>%
  filter(Round.Number == round_num) %>%
  select(-Time, -Round.Number) %>%
  mutate_at("Probability", formattable::percent, digits = 1) %>%
  formattable(list(
    Home.Team = formatter(
      "span", style =
        ~ style("font-weight" = ifelse(Probability >= 0.501, "bold", NA))
    ),
    Away.Team = formatter(
      "span", style =
        ~ style("font-weight" = ifelse(Probability <= 0.499, "bold", NA))
    )
  ))
```

## Ratings

```{r get data, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
# Load libraries
library(pacman)
pacman::p_load(fitzRoy, tidyverse, formattable, 
               ggthemes, elo, here, lubridate, ggrepel, 
               widgetframe, ggridges, DT, scales)

# Load data that has been run using 'weekly_data_process.R'
dat_list <- read_rds(here::here("data", "raw-data", "AFLM.rds"))
sim_dat_list <- read_rds(here::here("data", "raw-data", "AFLM_sims.rds"))
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
# Get the round data
elo_round <- dat_list$elo %>%
  filter(Game == last(Game) & !Team %in% c("Fitzroy", "University"))

sim_round <- sim_dat_list$sim_data_summary %>%
  filter(Season == max(Season)) %>%
  filter(Round == last(dat_list$elo$Round.Number))

# Combine Simulation and ELO data
elo_table <- elo_round %>%
  left_join(sim_round, by = "Team") %>%
  ungroup() %>%
  mutate(ELO.change = ELO - ELO_pre) %>%
  mutate_at(c("Top.8", "Top.4", "Top.1"), formattable::percent, digits = 1) %>%
  mutate_at(c("ELO", "ELO.change"), as.integer) %>%
  mutate_at("Wins", round, 1) %>%
  select(Team, ELO, ELO.change, Wins, Top.8, Top.4, Top.1) %>%
  arrange(desc(Wins))
```

```{r formattable, echo=FALSE, message=FALSE, warning=FALSE}
# Write to formattable with some formatting
dt_elo <- elo_table %>%
  formattable(list(
    ELO = normalize_bar("#20B2AA80"),
    ELO.change = formatter(
      "span",
      style = ~ style(color = ifelse(ELO.change >= 0, "green", "red"))
    ),
    Top.8 = color_tile("transparent", "lightblue"),
    Top.4 = color_tile("transparent", "lightblue"),
    Top.1 = color_tile("transparent", "lightblue")
  )) %>%
  as.datatable(options = list(dom = "t", pageLength = 18))

frameWidget(dt_elo, height = 700, width = "95%")
```

## Simulations

And below we show the probability of each team finishing at each ladder position in a heatmap. 
```{r sim_code, message=FALSE, warning=FALSE, include=FALSE}
sims <- 10000

# Get Table of percentages
simCount <- sim_dat_list$simCount %>%
    filter(Season == max(Season)) %>%
  filter(Round == max(Round))

## Reorder table by number of wins
# Find order of wins
simWins <- 
  sim_dat_list$sim_data_summary %>%
  filter(Season == max(Season)) %>%
  filter(Round == max(Round)) %>%
  arrange(Wins)

# Refactor
simCount$Team <- factor(simCount$Team, levels = simWins$Team)

# Get rankings within team
simCount <- simCount %>% 
  group_by(Team) %>% 
  mutate(order = dense_rank(desc(Freq)),
         txt = case_when(
           Freq < 1 ~ "<1", 
           Freq > 1 ~ as.character(round(Freq, 0)), 
           TRUE ~ "")) %>%
  arrange(Team, order)

# Define Quantiles
quant <- quantile(simCount$Freq, na.rm = T)

```

```{r sim_plotr, echo=FALSE, message=FALSE, warning=FALSE}
# Plot
ggplot(simCount, aes(x = Rank, y = Team)) +
  geom_tile(aes(fill = Freq)) +
  scale_fill_gradientn(colours = c("#ECECEC", "#ffffcc", "#bd0026"),
                       values = rescale(c(quant["0%"], quant["25%"], quant["100%"])),
                       na.value = "white") +
  geom_text(aes(label = txt, alpha = 0.5)) +
  scale_x_discrete(position = "top") +
  theme_minimal() + 
  guides(alpha = F, fill = F) + 
  labs(title = "AFL Mens Finishing Position",
       subtitle = "Probability of each team finishing in each ladder position",
       caption = "data sourced from afltables.com")

```