---
title: fitzRoy v1.1.0
author: James Day
date: '2022-01-11'
slug: []
categories:
  - fitzroy
tags: []
subtitle: ''
---

I'm excited to announce the official v1.1.0 release of fitzRoy has landed on CRAN!

This release includes a few new features and some improvements to bugs that had been reported. Read on to find out about them. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(fitzRoy)
library(tidyverse)
```
## Player Details
Newly included are the `player_details` functions. These follow the same pattern as the main `fetch_` family of functions, with some slight differences.

You can use `fetch_player_details` to return a list of player details for a given club and source. The source will determine exactly what details are returned. The default is the AFL.com.au website. 

The AFL website will return AFLW or AFLM data, while other sources only include AFLM data. 

```{r}
aflw_players <- fetch_player_details(team = "Western Bulldogs", current = TRUE, comp = "AFLW", source = "AFL")

head(aflw_players)
```

The list of details for AFL.com.au website is below. 

```{r}
glimpse(aflw_players)

```

The AFLTables.com source will return all players who have played for the specified team over time.

```{r}
fetch_player_details("Hawthorn", source = "afltables")
```

The Footywire.com source allows you to specify just the current list of players, which is fairly quick, or all players over time which can be a little slow. 

```{r}
fetch_player_details("Richmond", source = "footywire", current = TRUE)
```

One challenge is that each source accepts different values for the `team` argument. There is a helper function that will return an error if the wrong team is supplied and provide a list of acceptable values for that specific source.

You can read full details on the [Player Details Vignette](https://jimmyday12.github.io/fitzRoy/articles/player-details.html)


```{r, error=TRUE}
fetch_player_details("Greater Western Sydney", source = "afltables")
```

## Coaches votes
A new set of functions will return coaches votes. 

```{r echo=TRUE}
coaches <- fetch_coaches_votes(season = 2021, round_number = 1, team = "Hawthorn")
coaches
```

An associated function will also calculate possible ways in which the cochaes may have voted. 


```{r}
calculate_coaches_vote_possibilities(coaches, "Coach View")

```
You can read about these functions fully at the new [Vignette](https://jimmyday12.github.io/fitzRoy/articles/aflca-coaches-votes.html)

# Other changes
You can read about the rest of the bugs in the [changelog](https://jimmyday12.github.io/fitzRoy/news/index.html)